---
title: "Assignment6"
author: "Saurav Prem Kaushik Chetry"
date: "November 30, 2019"
output:
  html_document: default
  pdf_document: default
netid: schetry2@illinois.edu
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 1, digits = 3)
library(rjags)
library(lattice)
library(MASS)
library(knitr)
```

**Solution 1**  

```{r}
illini = read.csv("illinimensbb.csv")
head(illini)
plot(Ht~Pos, data = illini)
levels(illini$Pos)
unclass(illini$Pos)
```
    
  The players at center(C) positions had the largest heights. The median height was about 84.   
  The players at the forward(F) positions had the average heights. The median height was about 78.  
  The players at the guard(G) positions had the smallest heights.The median height was about 75.  
  The different field positions seem to prefer players with different heights.  
  
**Solution 2**  
  
**Part a**  
  
```{r eval=FALSE}

model {
for (i in 1:length(fgm)) {
fgm[i] ~ dbin(prob[i], fga[i])
logit(prob[i]) <- betapos[pos[i]] + betaht*htscaled[i]

fgmrep[i] ~ dbin(prob[i], fga[i])
}
  
for (j in 1:max(pos)) {
betapos[j] ~ dt(0, 0.01, 1)
}
betaht ~ dt(0, 0.16, 1)
}

```

```{r}
d1 <- list( fgm = illini$FGM,
            fga = illini$FGA,
            pos = unclass(illini$Pos),
            htscaled = as.vector(scale(illini$Ht, scale=2*sd(illini$Ht))))

```
 
```{r}
inits1 <- list(list(betapos=c(10,10,10), betaht=10),
               list(betapos=c(10,10,-10), betaht=-10),
               list(betapos=c(10,-10,10), betaht=-10),
               list(betapos=c(10,-10,-10), betaht=10))
```

```{r}
m1 <- jags.model("illini1.bug", d1, inits1, n.chains=4, n.adapt=1000)
```
  
  
```{r}
update(m1, 1000) # burn-in
```

```{r}
x1 <- coda.samples(m1, c("betapos","betaht"), n.iter=8000)
```

```{r}
gelman.diag(x1, autoburnin=FALSE)
```

```{r}
effectiveSize(x1)
```

**Part b**  

```{r}
summary(x1)
```

**Part c**  

```{r}
post.samp1 = as.matrix(x1)
betaht = post.samp1[,1]
betaposC = post.samp1[,2]
betaposF = post.samp1[,3]
betaposG = post.samp1[,4]

par(mfrow =c (1,3))
plot(betaht,betaposC)
plot(betaht,betaposF)
plot(betaht,betaposG)


```

**The scatterplots indicate linear relationship between the regression parametes. This indicates posterior correlations.**  

**Part d**  

```{r}

x1 <- coda.samples(m1, c("betapos","betaht","prob","fgmrep"),n.iter=2000)
probs <- as.matrix(x1)[, paste("prob[",1:nrow(illini),"]", sep="")]
hist(probs[,11], freq = FALSE)
```

**Part e** 
```{r}
mean(post.samp1[,"betapos[2]"] > post.samp1[,"betapos[3]"])

```
```{r}
mean(post.samp1[,"betapos[3]"] > post.samp1[,"betapos[2]"])
```

Bayes Factor:

```{r}
mean(post.samp1[,"betapos[2]"] > post.samp1[,"betapos[3]"])/mean(post.samp1[,"betapos[3]"] > post.samp1[,"betapos[2]"])
```

Data Evidence for $\beta_F$>$\beta_G$ vs. $\beta_G$>$\beta_F$ is: Strong

**Part f**  

```{r}
fgmrep <- as.matrix(x1)[, paste("fgmrep[",1:nrow(illini),"]", sep="")]
Tchi <- numeric(nrow(fgmrep))
Tchirep <- numeric(nrow(fgmrep))
for(s in 1:nrow(fgmrep)){
 Tchi[s] <- sum((illini$FGM - illini$FGA*probs[s,])^2 /
 (illini$FGA*probs[s,]*(1-probs[s,])))
 Tchirep[s] <- sum((fgmrep[s,] - illini$FGA*probs[s,])^2 /
 (illini$FGA*probs[s,]*(1-probs[s,])))
 }
```


```{r}
mean(Tchirep >= Tchi)
```

**Quite substantial evidence of a problem, probably overdispersion. We need a better model that takes account of this**  

**Part g**   

**i**  
**New JAGS to allow for overdispersion:**  
```{r eval= FALSE}
model {
for (i in 1:length(fgm)) {
fgm[i] ~ dbin(prob[i], fga[i])
logit(prob[i]) <- betapos[pos[i]] + betaht*htscaled[i] + epsilon[i]

epsilon[i] ~ dnorm(0,1/sigmaepsilon^2)
fgmrep[i] ~ dbin(prob[i], fga[i])
}
  
for (j in 1:max(pos)) {
betapos[j] ~ dt(0, 0.01, 1)
}
betaht ~ dt(0, 0.16, 1)

sigmaepsilon ~ dunif(0,10)
}
```

```{r}
inits2 <- list(list(betapos=c(10,10,10), betaht=10, sigmaepsilon=0.01),
               list(betapos=c(10,10,-10), betaht=-10, sigmaepsilon=9),
               list(betapos=c(10,-10,10), betaht=-10, sigmaepsilon=0.01),
               list(betapos=c(10,-10,-10), betaht=10, sigmaepsilon=9))

```

```{r}
m2 <- jags.model("illini2.bug", d1, inits2, n.chains=4, n.adapt=1000)
```

```{r}
update(m2, 8000) # burn-in
```

```{r}
x2 <- coda.samples(m2, c("betapos","betaht","sigmaepsilon"), n.iter=50000)
```

```{r}
gelman.diag(x2, autoburnin=FALSE)
```

```{r}
effectiveSize(x2[,c("betapos[1]","betapos[2]","betapos[3]","betaht","sigmaepsilon")])
```

**ii**  

```{r}
post.samp2 = as.matrix(x2)
hist(post.samp2[,5], freq = FALSE)
```

**iii(Repeat of Part 2.e)**  

```{r}
mean(post.samp2[,"betapos[2]"] > post.samp2[,"betapos[3]"])

```

```{r}
mean(post.samp2[,"betapos[3]"] > post.samp2[,"betapos[2]"])
```

Bayes Factor:

```{r}
mean(post.samp2[,"betapos[2]"] > post.samp2[,"betapos[3]"])/mean(post.samp2[,"betapos[3]"] > post.samp2[,"betapos[2]"])
```

Data Evidence for $\beta_F$>$\beta_G$ vs. $\beta_G$>$\beta_F$ is: Positive to Barely Mentionable. Previously the data evidence was Strong


**Solution 3**

**Part a**  

```{r, eval = FALSE}
model {
  for (i in 1:length(blk)) {
    blk[i] ~ dpois(lambda[i])
    log(lambda[i]) <- log(t[i]) + betapos2[pos2[i]] + betaht2*htscaled2[i]
    
    blkrep[i] ~ dpois(lambda[i])
  }
  
  for (j in 1:max(pos2)) {
    betapos2[j] ~ dnorm(0, 0.0001)
  }
  betaht2 ~ dnorm(0, 0.0001)
}
  

```

```{r}
d3 <- list( blk = illini$BLK,
            t = illini$MIN,
            pos2 = unclass(illini$Pos),
            htscaled2 = as.vector(scale(illini$Ht)))
```

```{r}
inits3 <- list(list(betapos2=c(90,90,90), betaht2=90),
               list(betapos2=c(90,90,0.01), betaht2=0.01),
               list(betapos2=c(90,0.01,90), betaht2=0.01),
               list(betapos2=c(90,0.01,0.01), betaht2=90))

```

```{r}
m3 <- jags.model("illini3.bug", d3, inits3, n.chains=4, n.adapt=1000)
```

```{r}
update(m3, 2000) # burn-in
```

```{r}
x3 <- coda.samples(m3, c("betapos2","betaht2"), n.iter=20000)
```

```{r}
gelman.diag(x3, autoburnin=FALSE)
```
```{r}
effectiveSize(x3[,c("betapos2[1]","betapos2[2]","betapos2[3]","betaht2")])    
```

**Part b**  

```{r}
(sm3 = summary(x3))
```

```{r}
x3_1 <- coda.samples(m3, c("betapos2","betaht2","blk","blkrep","lambda"), n.iter=2000)
```

**Part c**  

approximate 95% central posterior credible interval for this factor:

```{r}
factor = matrix(c(exp(sm3[[2]][1,1]),exp(sm3[[2]][1,5])),nrow = 1, ncol = 2, byrow = TRUE)
dimnames(factor) = list( 
   c("Factor"),      
   c("2.5%", "97.5%")) 
factor
```

**As the inverval carries positive values for the mean rate of blocking shots, greater height seems to be associated with greater rate of blocking shots**  


**Part d**  

```{r}
blk <- as.matrix(x3_1)[, paste("blk[",1:nrow(illini),"]", sep="")]
blkrep <- as.matrix(x3_1)[, paste("blkrep[",1:nrow(illini),"]", sep="")]
lambdas <- as.matrix(x3_1)[, paste("lambda[",1:nrow(illini),"]", sep="")]

Tchi_blk <- numeric(nrow(blkrep))
Tchirep_blk <- numeric(nrow(blkrep))
for(s in 1:nrow(blkrep)){
 Tchi_blk[s] <- sum((illini$BLK - lambdas[s,])^2 / lambdas[s,])
 Tchirep_blk[s] <- sum((blkrep[s,] - lambdas[s,])^2 / lambdas[s,])
}

mean(Tchirep_blk >= Tchi_blk)
```

**The posterior predictive p-value is suspiciously small.This could alternatively be an indicator of outliers.We conclude that there could be a bit of overdispersion**  

**Part e**  

**i**  

```{r warning=FALSE}
prob2 = rep(0,15)
for(i in 1:nrow(illini)){
  
 prob2[i] =  mean(blkrep[,i] >= illini$BLK[i])
 #prob2[i] =  mean(blkrep[,i] >= blk[,i]) #implies same prob calculation.
 
}
 (players = data.frame(illini$Player,prob2, illini$MIN, illini$BLK))
```


**ii**  

**Players with prov < 0.05:**    
```{r}
subset(players, players$prob2 < 0.05)
```

**iii**

**Players with prob = 1:**  
```{r}
subset(players, players$prob2 == 1)
```


**With the height and position relation in place, it is not surprising that a player would block lesser shots if he plays for short duration in the field. For such players who have the probability = 1, all have played for shorter time in the field and have blocked zero shots, as expected from the model**  









